\textbf{skater}: An R package for SNP-based Kinship Analysis, Testing, and Evaluation

### 2021-11-30 paper/paper.tex

\begin{abstract}
\hfill\break
\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation. A dedicated software package that consistently and intuitively implements this analysis functionality is needed.\\
\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}. The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.\\
\textbf{Availability:} The skater package is implemented as an R package and is released under the MIT license at https://github.com/signaturescience/skater. Documentation is available at https://signaturescience.github.io/skater.
\end{abstract}

@@

\textbf{R version}: R version 4.0.4 (2021-02-15)

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{install.packages}\NormalTok{(}\StringTok{"skater"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
\NormalTok{devtools}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"signaturescience/skater"}\NormalTok{, }\AttributeTok{build\_vignettes=}\ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(skater)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{famfile }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\AttributeTok{package=}\StringTok{"skater"}\NormalTok{, }\AttributeTok{mustWork=}\ConstantTok{TRUE}\NormalTok{)}
\NormalTok{fam }\OtherTok{\textless{}{-}} \FunctionTok{read\_fam}\NormalTok{(famfile)}
\NormalTok{fam}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds }\OtherTok{\textless{}{-}} \FunctionTok{fam2ped}\NormalTok{(fam)}
\NormalTok{peds}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(fid}\SpecialCharTok{==}\StringTok{"testped1"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 8 x 7
##   fid      id                dadid             momid        sex affected ped    
##   <chr>    <chr>             <chr>             <chr>      <int>    <dbl> <list> 
## 1 testped1 testped1_g1-b1-s1 <NA>              <NA>           1        1 <pedig~
## 2 testped1 testped1_g1-b1-i1 <NA>              <NA>           2        1 <pedig~
## 3 testped1 testped1_g2-b1-s1 <NA>              <NA>           1        1 <pedig~
## 4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 5 testped1 testped1_g2-b2-s1 <NA>              <NA>           1        1 <pedig~
## 6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_~     2        1 <pedig~
## 8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_~     1        1 <pedig~
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{mar=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\AttributeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_pedigree}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped, }\AttributeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ped2kinpair}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{kinpairs }\OtherTok{\textless{}{-}} 
\NormalTok{  peds }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{pairs=}\NormalTok{purrr}\SpecialCharTok{::}\FunctionTok{map}\NormalTok{(ped, ped2kinpair)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(fid, pairs) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{pairs)}
\NormalTok{kinpairs}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dibble}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dibble}\NormalTok{(}\AttributeTok{max\_degree =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Get two pairs from each type of relationship we have in kinpairs:}
\NormalTok{kinpairs\_subset }\OtherTok{\textless{}{-}} 
\NormalTok{  kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{group\_by}\NormalTok{(k) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{slice}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{)}
\NormalTok{kinpairs\_subset}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Infer degree out to third degree relatives:}
\NormalTok{kinpairs\_subset }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A \texttt{tibble} with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs.~the guessing rate.
\item
  A \texttt{tibble} with contingency table statistics calculated for each class. Details on the statistics calculated for each class can be reviewed on the help page for \texttt{?confusion\_matrix}.
\item
  A \texttt{matrix} with the contingency table object itself.
\item
  A \texttt{vector} with the reciprocal RMSE (R-RMSE). The R-RMSE is calculated as \texttt{sqrt(mean((1/(Target+.5)-1/(Predicted+.5))\^{}2)))}, and represents an alternative to classification accuracy when benchmarking relationship degree estimation. Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree). The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.
\end{enumerate}

To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Function to randomly flip levels of a factor (at 20\%, by default)}
\NormalTok{randomflip }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, }\AttributeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{runif}\NormalTok{(}\FunctionTok{length}\NormalTok{(x))}\SpecialCharTok{\textless{}}\NormalTok{p, }\FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(x)), x)}

\CommentTok{\# Infer degree (truth/target) using kin2degree, then randomly flip 20\% of them}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
\NormalTok{kinpairs\_inferred }\OtherTok{\textless{}{-}}\NormalTok{ kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\NormalTok{tidyr}\SpecialCharTok{::}\FunctionTok{replace\_na}\NormalTok{(degree\_truth, }\StringTok{"unrelated"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_inferred=}\FunctionTok{randomflip}\NormalTok{(degree\_truth))}
\NormalTok{kinpairs\_inferred}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 288 x 6
##    fid      id1               id2                   k degree_truth degree_inferred
##    <chr>    <chr>             <chr>             <dbl> <chr>        <chr>          
##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   0            0              
##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     unrelated    unrelated      
##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     unrelated    unrelated      
##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  1            1              
##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0     unrelated    unrelated      
##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  1            1              
##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 2            2              
##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 2            1              
##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5   0            0              
## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0     unrelated    unrelated      
## # ... with 278 more rows
\end{verbatim}

Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $Accuracy
## # A tibble: 1 x 5
##   Accuracy `Accuracy LL` `Accuracy UL` `Accuracy Guessing` `Accuracy P-value`
##      <dbl>         <dbl>         <dbl>               <dbl>              <dbl>
## 1    0.812         0.763         0.856               0.333           1.09e-62
## 
## $Other
## # A tibble: 6 x 15
##   Class         N `Sensitivity/Re~ `Specificity/TN~ `PPV/Precision`   NPV `F1/Dice`
##   <chr>     <dbl>            <dbl>            <dbl>           <dbl> <dbl>     <dbl>
## 1 0          64              0.75             0.964           0.857 0.931     0.8  
## 2 1          72              0.806            0.944           0.829 0.936     0.817
## 3 2          48              0.833            0.967           0.833 0.967     0.833
## 4 3           8              0.75             0.936           0.25  0.992     0.375
## 5 unrelated  96              0.854            0.958           0.911 0.929     0.882
## 6 Average    57.6            0.799            0.954           0.736 0.951     0.741
## # ... with 8 more variables: Prevalence <dbl>, Detection Rate <dbl>,
## #   Detection Prevalence <dbl>, Balanced Accuracy <dbl>, FDR <dbl>, FOR <dbl>,
## #   FPR/Fallout <dbl>, FNR <dbl>
## 
## $Table
##            Target
## Predicted    0  1  2  3 unrelated
##   0         48  4  2  1         1
##   1          5 58  4  0         3
##   2          0  3 40  1         4
##   3          8  4  0  6         6
##   unrelated  3  3  2  0        82
## 
## $recip_rmse
## [1] 0.4665971
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth) }\SpecialCharTok{\%\textgreater{}\%} 
\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hapibd\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
                               \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
\NormalTok{hapibd\_seg }\OtherTok{\textless{}{-}} \FunctionTok{read\_ibd}\NormalTok{(hapibd\_filepath, }\AttributeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
\NormalTok{hapibd\_seg}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ibd\_dat }\OtherTok{\textless{}{-}} \FunctionTok{ibd2kin}\NormalTok{(}\AttributeTok{.ibd\_data=}\NormalTok{hapibd\_seg, }\AttributeTok{.map=}\NormalTok{gmap)}
\NormalTok{ibd\_dat}
\end{Highlighting}
\end{Shaded}

>>>>

\begin{abstract}
\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation. A dedicated software package that consistently and intuitively implements this analysis functionality is needed.\\
\textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}. The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree package is implemented as an R package and is released under the MIT license at \textbf{Results:} Here we present the skater R package for \textbf{S}NP-based \textbf{k}inship classification, and summarizing IBD segment data.\\
\textbf{Availability:} The skater https://github.com/signaturescience/skater. Documentation is available at https://signaturescience.github.io/skater.
\end{abstract}

@@

\textbf{R version}: R version 4.0.2 (2020-06-22)

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"skater"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"signaturescience/skater"}\NormalTok{, }\DataTypeTok{build_vignettes=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(skater)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{famfile <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\DataTypeTok{package=}\StringTok{"skater"}\NormalTok{, }\DataTypeTok{mustWork=}\OtherTok{TRUE}\NormalTok{)}
\NormalTok{fam <-}\StringTok{ }\KeywordTok{read_fam}\NormalTok{(famfile)}
\NormalTok{fam}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds <-}\StringTok{ }\KeywordTok{fam2ped}\NormalTok{(fam)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: replacing previous import 'vctrs::data_frame' by 'tibble::data_frame'
## when loading 'dplyr'
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(fid}\OperatorTok{==}\StringTok{"testped1"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols=}\NormalTok{data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 8 x 7
##   fid      id                dadid             momid        sex affected ped    
##   <chr>    <chr>             <chr>             <chr>      <int>    <dbl> <list> 
## 1 testped1 testped1_g1-b1-s1 NA                NA             1        1 <pedig~
## 2 testped1 testped1_g1-b1-i1 NA                NA             2        1 <pedig~
## 3 testped1 testped1_g2-b1-s1 NA                NA             1        1 <pedig~
## 4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 5 testped1 testped1_g2-b2-s1 NA                NA             1        1 <pedig~
## 6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
## 7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_~     2        1 <pedig~
## 8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_~     1        1 <pedig~
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\DataTypeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot_pedigree}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped, }\DataTypeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ped2kinpair}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{kinpairs <-}\StringTok{ }
\StringTok{  }\NormalTok{peds }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pairs=}\NormalTok{purrr}\OperatorTok{::}\KeywordTok{map}\NormalTok{(ped, ped2kinpair)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(fid, pairs) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols=}\NormalTok{pairs)}
\NormalTok{kinpairs}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dibble}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dibble}\NormalTok{(}\DataTypeTok{max_degree =} \DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get two pairs from each type of relationship we have in kinpairs:}
\NormalTok{kinpairs_subset <-}\StringTok{ }
\StringTok{  }\NormalTok{kinpairs }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(k) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{slice}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{)}
\NormalTok{kinpairs_subset}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Infer degree out to third degree relatives:}
\NormalTok{kinpairs_subset }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree=}\KeywordTok{kin2degree}\NormalTok{(k, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A \texttt{tibble} with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs.~the guessing rate.
\item
  A \texttt{tibble} with contingency table statistics calculated for each class. Details on the statistics calculated for each class can be reviewed on the help page for \texttt{?confusion\_matrix}.
\item
  A \texttt{matrix} with the contingency table object itself.
\item
  A \texttt{vector} with the reciprocal RMSE (R-RMSE). The R-RMSE represents an alternative to classification accuracy when benchmarking relationship degree estimation and is calculated using the formula in (1). Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree). The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.
\end{enumerate}

\begin{equation}
\sqrt{\frac{\sum_{i=1}^{k}(\frac{1}{\text{Target}+0.5}-\frac{1}{\text{Predicted}+0.5})^2}{k}}
\end{equation}

To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Function to randomly flip levels of a factor (at 20%, by default)}
\NormalTok{randomflip <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x, }\DataTypeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{runif}\NormalTok{(}\KeywordTok{length}\NormalTok{(x))}\OperatorTok{<}\NormalTok{p, }\KeywordTok{sample}\NormalTok{(}\KeywordTok{unique}\NormalTok{(x)), x)}

\CommentTok{# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
\NormalTok{kinpairs_inferred <-}\StringTok{ }\NormalTok{kinpairs }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_truth=}\KeywordTok{kin2degree}\NormalTok{(k, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_truth=}\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{replace_na}\NormalTok{(degree_truth, }\StringTok{"unrelated"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_inferred=}\KeywordTok{randomflip}\NormalTok{(degree_truth))}
\NormalTok{kinpairs_inferred}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 288 x 6
##    fid      id1               id2                 k degree_truth degree_inferred
##    <chr>    <chr>             <chr>           <dbl> <chr>        <chr>          
##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1~ 0.5   0            0              
##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1~ 0     unrelated    unrelated      
##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1~ 0     unrelated    unrelated      
##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1~ 0.25  1            1              
##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2~ 0     unrelated    unrelated      
##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2~ 0.25  1            1              
##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1~ 0.125 2            2              
##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2~ 0.125 2            1              
##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1~ 0.5   0            0              
## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1~ 0     unrelated    unrelated      
## # ... with 278 more rows
\end{verbatim}

Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{confusion_matrix}\NormalTok{(}\DataTypeTok{prediction =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_inferred, }
                 \DataTypeTok{target =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_truth)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $Accuracy
## # A tibble: 1 x 5
##   Accuracy `Accuracy LL` `Accuracy UL` `Accuracy Guessing` `Accuracy P-value`
##      <dbl>         <dbl>         <dbl>               <dbl>              <dbl>
## 1    0.812         0.763         0.856               0.333           1.09e-62
## 
## $Other
## # A tibble: 6 x 15
##   Class      N `Sensitivity/Re~ `Specificity/TN~ `PPV/Precision`   NPV `F1/Dice`
##   <chr>  <dbl>            <dbl>            <dbl>           <dbl> <dbl>     <dbl>
## 1 0       64              0.75             0.964           0.857 0.931     0.8  
## 2 1       72              0.806            0.944           0.829 0.936     0.817
## 3 2       48              0.833            0.967           0.833 0.967     0.833
## 4 3        8              0.75             0.936           0.25  0.992     0.375
## 5 unrel~  96              0.854            0.958           0.911 0.929     0.882
## 6 Avera~  57.6            0.799            0.954           0.736 0.951     0.741
## # ... with 8 more variables: Prevalence <dbl>, Detection Rate <dbl>,
## #   Detection Prevalence <dbl>, Balanced Accuracy <dbl>, FDR <dbl>, FOR <dbl>,
## #   FPR/Fallout <dbl>, FNR <dbl>
## 
## $Table
##            Target
## Predicted    0  1  2  3 unrelated
##   0         48  4  2  1         1
##   1          5 58  4  0         3
##   2          0  3 40  1         4
##   3          8  4  0  6         6
##   unrelated  3  3  2  0        82
## 
## $recip_rmse
## [1] 0.4665971
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{confusion_matrix}\NormalTok{(}\DataTypeTok{prediction =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_inferred, }
                 \DataTypeTok{target =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_truth) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{purrr}\OperatorTok{::}\KeywordTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hapibd_filepath <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
                               \DataTypeTok{package=}\StringTok{"skater"}\NormalTok{)}
\NormalTok{hapibd_seg <-}\StringTok{ }\KeywordTok{read_ibd}\NormalTok{(hapibd_filepath, }\DataTypeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
\NormalTok{hapibd_seg}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ibd_dat <-}\StringTok{ }\KeywordTok{ibd2kin}\NormalTok{(}\DataTypeTok{.ibd_data=}\NormalTok{hapibd_seg, }\DataTypeTok{.map=}\NormalTok{gmap)}
\NormalTok{ibd_dat}
\end{Highlighting}
\end{Shaded}
